---
output: 
  bookdown::pdf_document2:
    toc: False
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lastpage}
- \usepackage{amsmath}
- \usepackage{tocloft}
bibliography: ref.bib
link-citations: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(GGally)
library(dplyr)
library(bookdown)
```


```{python include=FALSE}
import numpy as np
import pandas as pd
import seaborn as sns
sns.set_style('whitegrid')
import matplotlib.pyplot as plt
import os
import time
from plotly import tools
from plotly.offline import plot
import plotly.offline as py
from plotly.graph_objs import Scatter, Layout
import plotly.graph_objs as go
import plotly.figure_factory as ff
```

\fancyfoot[C]{\thepage\ of \pageref{LastPage}}

\newpage
\begin{titlepage}
\begin{center}
\vspace*{1cm}
\large{STAT 6500}\\
\vspace*{2cm}
\line(1,0){400}\\
\huge{\textbf{Statistical Machine Learning}}\\
\vspace*{1cm}
Land Use Cover\\
\vspace*{1cm}
\normalsize{\textbf{Kendall Byrd - Atitarn Dechasuravanit - Alexys Rodriguez - Abdulaziz Alsugair}}\\

\vspace*{1cm}
\normalsize{Project Proposal}\\

\line(1,0){400}
\vfill
\huge{Spring 2022}\\
Wednesday, March 2
\end{center}
\end{titlepage}


\newpage
\tableofcontents
 

\newpage
\pagestyle{fancy}
\fancyhead[L]{\slshape Land Use Cover}
\fancyhead[C]{\slshape STAT 6500}
\fancyhead[R]{\slshape Project Proposal}

# EDA

```{r include=FALSE}
#Load and view Training data
trainData <- read.csv("training.csv")
trainClass = table(trainData$class)

#Load and view Test data
testData = read.csv('testing.csv')
testClass = table(testData$class)

Class = rbind(trainClass, testClass)

# convert class from "char" into a factor
trainData$class <- as.factor(trainData$class)

# convert class from "char" into a factor
testData$class <- as.factor(testData$class)

```

From the source @urbanlandcover the dataset is divided into training (168 instances) and testing (507 instances). The number of attributes for both datasets is 148. The first attribute is `class` which contains the target (y) variable and is detailed (training and testing) in Table \@ref(tab:class), and in Figure \@ref(fig:class2) for training data.


```{r class, echo=FALSE}
knitr::kable(Class, caption = "Target Attribute 'class'")
```

```{python class2, echo=FALSE, fig.align="center", fig.cap="Attribute 'class' in training", message=FALSE, warning=FALSE, fig.height=2.5, fig.width=5}
#load in training and test data
train = pd.read_csv('training.csv')

# check for missing values although it is clear there are none
train.isnull().any().any()

# duplicated function of pandas returns a duplicate row as true and others as false
sum(train.duplicated())
f,axes=plt.subplots(1,2,figsize=(20,8))
train['class'].value_counts().plot.pie(autopct='%1.1f%%',ax=axes[0])
axes[0].set_title('Visual for Distribution of Different Classes')
axes[0].set_ylabel('')
sns.countplot('class',data=train,ax=axes[1]) # sns.countplot is used
# like a histogram but for
# categorical data
axes[1].set_title('Visual for Distribution of Different Classes')
plt.show()

```




# References

